### 2010-10-14 papers/BMD2010/PhysicalParametersPaper/paper.tex
more paper work

The moments and products of inertia of the rear frame and handlebar/fork
assembly with reference to the benchmark coordinate system were calculated by
formulating the relationship between inertial frames
\begin{equation}
	\mathbf{J}_i=\mathbf{R}_i^T\mathbf{IR}_i
\label{eq:rotIn}
\end{equation}
where $\mathbf{J}_i$ is the inertia tensor about the pendulum axes,
$\mathbf{I}$, is the inertia tensor in the global reference frame and
$\mathbf{R}$ is the rotation matrix relating the two frames. The global inertia
tensor is defined as
\begin{equation}
	\mathbf{I}=
	\left[
	\begin{array}{rr}
		I_{xx}  & -I_{xz}\\
		-I_{xz} & I_{zz}
	\end{array}
	\right]\textrm{.}
	\label{eq:MoI}
\end{equation}
The inertia tensor can be reduced to a $2\times2$ matrix because the frame is
assumed to be laterally symmetric and the $y$ axis of the pendulum reference is
the same as the $y$ axis of the benchmark reference frame. The simple rotation matrix about the $Y$-axis
can similarly be reduced to a $2\times2$ matrix where $s_{\beta i}$ and
$c_{\beta i}$ are defined as $\sin{\beta_i}$ and $\cos{\beta_i}$,
respectively.
\begin{equation}
	\mathbf{R}=
	\left[
	\begin{array}{rr}
		c_{\beta i} & s_{\beta i}\\
		-s_{\beta i} & c_{\beta i}
	\end{array}
	\right]
	\label{eq:rotMat}
\end{equation}
The first entry of $\mathbf{J}_i$ in Eq.~\ref{eq:rotIn} is the moment of inertia about the pendulum axis and is written explicitly as 
\begin{equation}
	J_{i}=c^{2}_{\beta i}I_{xx}+2s_{\beta i}c_{\beta i}I_{xz}+s^{2}_{\beta i}I_{zz}\textrm{.}
\label{eq:inRelComp}
\end{equation}
Calculating all three $J_{i}$ allows one to form
\begin{equation}
	\left[
	\begin{array}{c}
		J_{1}\\
		J_{2}\\
		J_{3}
	\end{array}
	\right]
	=
	\left[
	\begin{array}{ccc}
		c_{\beta 1}^2 & -2s_{\beta 1}c_{\beta 1} & s_{\beta 1}^2\\
		c_{\beta 2}^2 & -2s_{\beta 2}c_{\beta 2} & s_{\beta 2}^2\\
		c_{\beta 3}^2 & -2s_{\beta 3}c_{\beta 3} & s_{\beta 3}^2
	\end{array}
	\right]
	\left[
	\begin{array}{c}
		I_{xx}\\
		I_{xz}\\
		I_{zz}
	\end{array}
	\right]
\label{eq:inRel}
\end{equation}
and the unknown global inertia tensor can be solved for.
\begin{figure}[tbp]
    \centering
    \subfloat[]{
        \label{fig:frameCompound}
        \includegraphics[width=2.75in]{../../../images/YellowFrameCompoundFirst.jpg}
        }
    \subfloat[]{
        \label{fig:forkCompound}
        \includegraphics[width=2.75in]{../../../images/BrowserInsForkCompoundFirst.jpg}
        }
        \caption{\subref{fig:frameCompound} The rear frame hung as a compound
        pendulum. \subref{fig:forkCompound} The Browser fork hung as a
        compound pendulum.}
        \label{fig:compound}
\end{figure}
\subsection{Fork}
\begin{figure}[tbp]
    \centering
    \subfloat[]{
        \label{fig:StratosFork}
        \includegraphics[width=2.75in]{../../../images/StratosForkTorsionalThird.jpg}
        }
    \caption{\subref{fig:StratosFork} The Stratos fork and handlebar assembly hung as a torsional
        pendulum.}
    \label{fig:torsional}
\end{figure}
\section{LINEAR ANALYSIS}
Once all of the bicycle parameters have been calculated the canonical matrices
can be formed and the linear dynamics of the bicycles can be explored. We also
added the same rigid rider to each bicycle for further comparison.
\begin{table}[tbp]
    \centering
    \caption{Rider parameters}
    \begin{tabular}{ll}
        Parameter & Value\\
        \hline
        $m_\mathrm{P}$ [kg] & 72\\
        $x_\mathrm{P}$ [m] & 0.2909\\
        $z_\mathrm{P}$ [m] & -1.1091\\
        $I_\mathrm{P}$ [$\mathrm{kg\ m}^2$] &
            $\left[
            \begin{array}{ccc}
                 7.9985 & 0      & -1.9272\\
                 0      & 8.0689 & 0\\
                -1.9272 & 0      & 2.3624
            \end{array}
            \right]$
    \end{tabular}
\end{table}

\subsection{EIGENVALUES}
\begin{figure}[tbp]
    \centering
    \includegraphics[width=4in]{../../../plots/Bike/eig_plot.pdf}
    \caption{The eigenvalues verus speed for all eight bicycles.}
    \label{fig:bikeEigPlot}
\end{figure}
\begin{figure}[tbp]
    \begin{center}
        \includegraphics[width=4in]{../../../plots/Bike/CrescendoRootLoci.pdf}
    \end{center}
    \caption{The root loci with speed as the parameter for the Crescendo.}
    \label{fig:rootloci}
\end{figure}
\begin{figure}[tbp]
    \centering
    \includegraphics[width=4in]{../../../plots/BikeRider/eig_plot.pdf}
    \caption{The eigenvalues verus speed for all eight bicycles with the same
    rider.}
    \label{fig:bikeRiderEigPlot}
\end{figure}
\subsection{FREQUENCY RESPONSE}
\bibliographystyle{acm}
\bibliography{bicycle}
\appendix
\section{PARAMETERS}
\subsection{BICYCLE}
\begin{landscape}
\subfile{../../../tables/Bike/Parameters/MasterParTable.tex}
\end{landscape}
\subsection{BICYCLE WITH RIDER}
%\begin{landscape}
%\subfile{../../../tables/Bike/Parameters/MasterParTable.tex}
%\end{landscape}
\section{CANONICAL MATRICES}
\subsection{BICYCLE}
\begin{landscape}
\subfile{../../../tables/Bike/Canonical/MasterCanTable.tex}
\end{landscape}
\subsection{BICYCLE WITH RIDER}
%\begin{landscape}
%\subfile{../../../tables/Bike/Parameters/MasterParTable.tex}
%\end{landscape}
\end{document}

>>>>

The moments and products of inertia of the rear frame and handlebar/fork
assembly with reference to the benchmark coordinate system were calculated by
formulating the relationship between inertial frames
	\mathbf{J}_i=\mathbf{R}_i^T\mathbf{IR}_i
\label{eq:rotIn}
where $\mathbf{J}_i$ is the inertia tensor about the pendulum axes,
$\mathbf{I}$, is the inertia tensor in the global reference frame and
$\mathbf{R}$ is the rotation matrix relating the two frames. The global inertia
tensor is defined as
	\mathbf{I}=
	\begin{array}{rr}
		I_{xx}  & -I_{xz}\\
		-I_{xz} & I_{zz}
	\right]\textrm{.}
	\label{eq:MoI}
The inertia tensor can be reduced to a $2\times2$ matrix because the frame is
assumed to be laterally symmetric and the $y$ axis of the pendulum reference is
the same as the $y$ axis of the benchmark reference frame. The simple rotation matrix about the $Y$-axis
can similarly be reduced to a $2\times2$ matrix where $s_{\beta i}$ and
$c_{\beta i}$ are defined as $\sin{\beta_i}$ and $\cos{\beta_i}$,
respectively.
	\mathbf{R}=
	\begin{array}{rr}
		c_{\beta i} & s_{\beta i}\\
		-s_{\beta i} & c_{\beta i}
	\label{eq:rotMat}
The first entry of $\mathbf{J}_i$ in Eq.~\ref{eq:rotIn} is the moment of inertia about the pendulum axis and is written explicitly as 
	J_{i}=c^{2}_{\beta i}I_{xx}+2s_{\beta i}c_{\beta i}I_{xz}+s^{2}_{\beta i}I_{zz}\textrm{.}
\label{eq:inRelComp}
Calculating all three $J_{i}$ allows one to form
		J_{1}\\
		J_{2}\\
		J_{3}
	\begin{array}{ccc}
		c_{\beta 1}^2 & -2s_{\beta 1}c_{\beta 1} & s_{\beta 1}^2\\
		c_{\beta 2}^2 & -2s_{\beta 2}c_{\beta 2} & s_{\beta 2}^2\\
		c_{\beta 3}^2 & -2s_{\beta 3}c_{\beta 3} & s_{\beta 3}^2
		I_{xx}\\
		I_{xz}\\
		I_{zz}
\label{eq:inRel}
and the unknown global inertia tensor can be solved for.
    \subfloat[]{
        \label{fig:frameCompound}
        \includegraphics[width=2.75in]{../../../images/YellowFrameCompoundFirst.jpg}
    \subfloat[]{
        \label{fig:forkCompound}
        \includegraphics[width=2.75in]{../../../images/BrowserInsForkCompoundFirst.jpg}
        \caption{\subref{fig:frameCompound} The rear frame hung as a compound
        pendulum. \subref{fig:forkCompound} The Browser fork hung as a
        compound pendulum.}
        \label{fig:compound}
    \subfloat[]{
        \label{fig:StratosFork}
        \includegraphics[width=2.75in]{../../../images/StratosForkTorsionalThird.jpg}
    \caption{\subref{fig:StratosFork} The Stratos fork and handlebar assembly hung as a torsional
        pendulum.}
    \label{fig:torsional}
\section{LINEAR ANALYSIS}
Once all of the bicycle parameters have been calculated the canonical matrices
can be formed and the linear dynamics of the bicycles can be explored. We also
added the same rigid rider to each bicycle for further comparison. The rigid
rider was assumed to be in same position and posture for each bicycle relative
to the rear wheel contact point.
\begin{table}[tbp]
    \begin{tabular}{ll}
        Parameter & Value\\
        \hline
        $m_\mathrm{P}$ [kg] & 72\\
        $x_\mathrm{P}$ [m] & 0.2909\\
        $z_\mathrm{P}$ [m] & -1.1091\\
        $I_\mathrm{P}$ [$\mathrm{kg\ m}^2$] &
            $\left[
            \begin{array}{ccc}
                 7.9985 & 0      & -1.9272\\
                 0      & 8.0689 & 0\\
                -1.9272 & 0      & 2.3624
            \right]$
    \end{tabular}
    \caption{Mass, center of mass and moment of inertia for the Person relative
    to the benchmark coordinate system from~\cite{Moore2009a}}
    \label{tab:riderParam}
\end{table}
\subsection{EIGENVALUES}
    \includegraphics[width=4in]{../../../plots/Bike/eig_plot.pdf}
    \caption{The eigenvalues verus speed for all eight bicycles.}
    \label{fig:bikeEigPlot}
The eigenvalues can be plotted versus forward speed.
Figure~\ref{fig:bikeEigPlot} shows that the bikes have the typical
characteristics of the Whipple bicycle model: four real roots at very slow
speeds, two of which are unstable; a complex pair that is unstable and lower
speeds and stable at higher speeds; and a root that is sligthly unstable at
higher speeds. The one noticable difference is that the capsize and caster
modes colasece into a complex pair between about 0.5 and 3 m/s. The frequency
of oscillation is similar to that of the weave mode. The root locus in the real
and imaginary plane, Fig.~\ref{fig:rootloci}, shows that the mode damps out
quickly. Examining the eigenvectors reveals that the mode is steer leading
roll with a 90 degree phase, both of their magitudes being similar,
Fig~\ref{fig:evec}.
        \includegraphics[width=4in]{../../../plots/Bike/CrescendoRootLoci.pdf}
    \caption{The root loci with speed as the parameter for the Crescendo.}
    \label{fig:rootloci}
        \includegraphics[width=3in]{../../../plots/cres.pdf}
    \caption{The root loci with speed as the parameter for the Crescendo.}
    \label{fig:evec}
With the rider added, the second complex pair disappears and the bikes have the
typical characteristics of the Whipple model. Reversing the fork on the yellow
bike lowers the weave critical speed. Also the addition of weight to the rear
rack of the Browser does blank.
    \includegraphics[width=4in]{../../../plots/BikeRider/eig_plot.pdf}
    \caption{The eigenvalues verus speed for all eight bicycles with the same
    rider.}
    \label{fig:bikeRiderEigPlot}
\subsection{FREQUENCY RESPONSE}
The frequency response of the bicycles and bicycle with rider also reveal some
interesting things. In the steer torque to roll Bode diagram the magnitude
diffence among bicycles can vary up to 10 dB (0.148 radians per newton meter of
torque).
    \includegraphics[width=4in]{../../../plots/Bike/Bode/Tdel2phi.pdf}
    \caption{The eigenvalues verus speed for all eight bicycles with the same
    rider.}
    \label{fig:bikeBode}
    \includegraphics[width=4in]{../../../plots/BikeRider/Bode/Tdel2phi.pdf}
    \caption{The eigenvalues verus speed for all eight bicycles with the same
    rider.}
    \label{fig:bikeRiderBode}

\bibliographystyle{acm}
\bibliography{bicycle}
\appendix
\section{RAW DATA}
%\begin{landscape}
%\subfile{../../../tables/Bike/Parameters/MasterParTable.tex}
%\end{landscape}
\section{PARAMETERS}
\subsection{BICYCLE}
\begin{landscape}
\subfile{../../../tables/Bike/Parameters/MasterParTable.tex}
\end{landscape}
\subsection{BICYCLE WITH RIDER}
%\begin{landscape}
%\subfile{../../../tables/Bike/Parameters/MasterParTable.tex}
%\end{landscape}
\section{CANONICAL MATRICES}
\subsection{BICYCLE}
\begin{landscape}
\subfile{../../../tables/Bike/Canonical/MasterCanTable.tex}
\end{landscape}
\subsection{BICYCLE WITH RIDER}
%\begin{landscape}
%\subfile{../../../tables/Bike/Parameters/MasterParTable.tex}
%\end{landscape}
\end{document}
